### 2021-09-18 plos_latex_template.tex
Rocío: Minor edits to rule 2 inclusion

Investing more in effort in the most excluded groups does not mean neglecting the others (see Rule 2 for more on that). 
But it will prevent from falling into the mistake of treating diversity as a checklist.
It will also guide the vision of diversity for your conference--and your strategies to achieve it:
Will a more diverse conference would translate into an even gender distribution in your speakers? 
Would it be representation of all races--especially Black people--among the head organizers, speakers, and attendees? 
Would it be having LGBTQIA+ friendly-spaces or community participation from key geographic regions?
This exercise will also help coming up with indicators that would help the organizing team assess the fulfillment of the diversity objectives along the way. 
%sara: reference for privilege: McIntosh, Peggy. "White privilege: Unpacking the invisible knapsack." (1988).

@@

While it is essential to improve representation towards some of the most visible dimensions of human diversity, such as race, gender, and country of origin, building a truly inclusive environment means taking care of all the other aspects of diversity as well. Having consideration of religious practices, setting specific accommodations for breastfeeding and child care, having LGBTQIA+-friendly spaces, creating community-only spaces, enforcing the use of pronouns, and acknowledging gender is not binary are just some examples of decisions that can make inclusion real.
Intersectionality also needs to be taken into account, to address multiple inequalities and barriers to unprivileged people that intersect and interact with each other \cite{crenshaw1989}. 
%sara: i'd move the intersectionality phrase to the last paragraph of the first rule, right before 'This vision should guide and help hold the organizing team accountable along the way.' Here, I'd just write: 
Importantly, you can take active steps in creating a more welcoming environment without requiring anyone to disclose personal information.

Paying attention to some dimensions of diversity while neglecting others may have unintended, even harmful consequences. Lack of representation, unwelcoming --or overtly aggressive-- environments hinder participation (or future participation) of people who could otherwise become active community members. A negative conference experience can divert career paths, affect lives, and exclude people from some fields. %Rocío: it would be nice to have reference here.
%ast: maybe https://link.springer.com/article/10.1007/s11199-017-0800-9
%Adithi: Assist newbies of the conference navigate through the conference as they might feel overwhelmed attending a conference.

%i took the networking part to expand communication att.ast we can switch back afterwards

Adopting a code of conduct and creating a team to enforce it are key aspects in creating a safe environment during a conference \cite{favaroYourScienceConference2016}. The code of conduct is a document meant to keep the community safe and should state clearly: the unacceptable behaviors, the spaces of the conference in which it applies, the consequences for engaging in unacceptable behavior, and the way to report violations \cite{aurora2019}. 
The code of conduct should be displayed prominently in several spaces of the conference to deter people from unacceptable behavior.
%Rocío: Microagression. I know it's not always evident, but that's a detail I would expect in a CoC, not necessarily here unless explicitly justified.
An efficient code of conduct acts as a protection for the community because the people who are the target of unacceptable behavior tend to be the ones with less power or privilege.

The code of conduct team should receive training on how to receive reports, respond to incidents, and communicate their responses. A diverse code of conduct team will be more understanding of power dynamics and sensitive to discrimination and harassment issues. There are also disciplinary cultures and geographical considerations that may need to be considered when developing and communicating the code of conduct (CoC). We strongly recommend reading “How to Respond to Code of Conduct Reports” \cite{aurora2019} as an excellent starting point for the Code of Conduct team work. A Code of Conduct response committee needs to be trained and prepared in advance on how to decide on the incident reports.

@@

%Some teams gather this vision in a diversity statement (e.g., \url{https://user2021.r-project.org/about/diversity_statement/}) to express the values of the conference and make diversity part of the outreach and communication strategy see Rule \ref{rule_communication}. However, it is important to not stop with a statement alone, but to identify specific goals and tasks that ensure compliance with these broader goals, for example, identifying accessible platforms, developing a detailed code of conduct and a team to enforce it, and preparing accessibility guidelines (see other Rules below).

>>>>

Investing more in effort in the most excluded groups does not mean neglecting the others (see Rule 2 \ref{rule_inclusion} for more on that). 
But it will prevent from falling into the mistake of treating diversity as a checklist.
It will also guide the vision of diversity for your conference--and your strategies to achieve it:
Will a more diverse conference would translate into an even gender distribution in your speakers? 
Would it be representation of all races--especially Black people--among the head organizers, speakers, and attendees? 
Would it be having LGBTQIA+ friendly-spaces or community participation from key geographic regions?
This exercise will also help coming up with indicators that would help the organizing team assess the fulfillment of the diversity objectives along the way. 
%sara: reference for privilege: McIntosh, Peggy. "White privilege: Unpacking the invisible knapsack." (1988).

@@

While it is essential to improve representation towards some of the most visible dimensions of human diversity, such as race, gender, and country of origin, building a truly inclusive environment means taking care of all the other aspects of diversity as well. Having consideration of religious practices, setting specific accommodations for breastfeeding and child care, having LGBTQIA+-friendly spaces, creating community-only spaces, enforcing the use of pronouns, and acknowledging that gender is not binary are just some examples of decisions that can make inclusion real.
Importantly, you can take active steps in creating a more welcoming environment without requiring anyone to disclose personal information.

Paying attention to some dimensions of diversity while neglecting others may have unintended, even harmful consequences. Lack of representation, unwelcoming --or overtly aggressive-- environments hinder participation (or future participation) of people who could otherwise become active community members. A negative conference experience can divert career paths, affect lives, and exclude people from some fields. %Rocío: it would be nice to have reference here.
%ast: maybe https://link.springer.com/article/10.1007/s11199-017-0800-9
%Adithi: Assist newbies of the conference navigate through the conference as they might feel overwhelmed attending a conference.
% Rocío: Andrea, could the newbies part go in the networking (box)? I don't see how to make it fit here. 
%i took the networking part to expand communication att.ast we can switch back afterwards

Adopting a code of conduct and creating a team to enforce it are key aspects in creating a safe environment during a conference \cite{favaroYourScienceConference2016}.
The code of conduct is a document meant to keep the community safe and should state clearly: the unacceptable behaviors, the spaces of the conference in which it applies, the consequences for engaging in unacceptable behavior, and the way to report violations \cite{aurora2019}. 
The code of conduct should be displayed prominently in several spaces of the conference to deter people from unacceptable behavior.
An efficient code of conduct acts as a protection for the community because the people who are the target of unacceptable behavior tend to be the ones with less power or privilege.

The code of conduct team should receive training on how to receive reports, respond to incidents, and communicate their responses, and organize accordingly. A diverse code of conduct team will be more understanding of power dynamics and sensitive to discrimination and harassment issues. There are also disciplinary cultures and geographical considerations that may need to be considered when developing and communicating the code of conduct (CoC). We strongly recommend reading “How to Respond to Code of Conduct Reports” \cite{aurora2019} as an excellent starting point for the Code of Conduct team work.

@@

%Some teams gather this vision in a diversity statement (e.g., \url{https://user2021.r-project.org/about/diversity_statement/}) to express the values of the conference and make diversity part of the outreach and communication strategy see Rule \ref{rule_communication}. However, it is important to not stop with a statement alone, but to identify specific goals and tasks that ensure compliance with these broader goals, for example, identifying accessible platforms, developing a detailed code of conduct and a team to enforce it, and preparing accessibility guidelines (see other Rules below).
% Rocío: This could be moved to communication (Batool, Andrea)
